### 2014-09-22 Paper_v1.tex
Fig 9 sensitivity

The model ensemble predicted a direct correlation between thrombin generation and rFVIIa addition in hemophilia (Fig. \ref{fig-rFVII}).
In the current proof of concept model, we cannot distinguish between different initiation sources, e.g., TF/FVIIa and rFVIIa, as we have only a single lumped initiation source (trigger).
Thus, we simulated the addition of rFVIIa in hemophilia by removing fVIII and fIX from the model, and modulating the level of trigger.
The baseline level of trigger is consistent with the previous hemophilia simulations, where the only thrombin being produced is from initiation mechanisms  (Fig. \ref{fig-rFVII}, 1$\times$~trigger).
As we increased the strength of the trigger event, the thrombin peak time and the maximum value of thrombin increased (Fig. \ref{fig-rFVII}, 50$\times$~trigger). 
However, as the trigger strength increased, the thrombin generated quickly decayed.
For example, for the highest trigger strength simulated (50$\times$trigger), 95\% of the thrombin was gone by 20 min after initiation.

We performed flux analysis to understand how the reduced order coagulation model balanced initiation, amplification and inhibition of thrombin formation for normal coagulation and hemophilia.
Analysis of the reaction flux through the reduced order network for thrombin generation in normal, hemophilia and rFVIIa-treated hemophilia identified three distinct operational modes (Fig. \ref{fig-flux-analysis}).
The reduced order network includes four lumped reactions, initiation, amplification, thrombin-induced APC generation and total thrombin inhibition (including both APC and ATIII action).
Directly after the addition of a trigger (e.g., TF/FVIIa or rFVIIa), the lumped initiation flux was the largest for all three cases.
However, within a few minutes enough thrombin was generated by the initiation mechanism to induce the amplification stage. 
During amplification, thrombin catalyzes its own formation and inhibition by generating activated protein C (APC), a potent inhibitor of the coagulation cascade.
For normal coagulation, amplification and thrombin inhibition are the dominate reactions by 6 min after initiation (Fig. \ref{fig-flux-analysis}, left). 
After 10 min, the dominate reaction has shifted to thrombin inhibition (both ATIII and APC action). In the current proof-of-principle model, 
APC inhibits upstream of amplification thus APC activity will slow the rate of thrombin formation directly from the initiation trigger. 
On the other hand, ATIII inhibits thrombin directly, as well as upstream coagulation factors.
In hemophilia (missing both fVIII and fIX), the amplification reaction does not occur and the only thrombin produced is from initiation (Fig. \ref{fig-flux-analysis}, center).
Initiation is quickly inhibited by APC, and the thrombin level stabilizes and eventually decays because of ATIII activity. 
Lastly, when 50$\times$rFVIIa is used to induce thrombin formation in hemophilia (absence of fVIII/fIX), 
initiation mechanisms dominate for up to 6 min following initiation (Fig. \ref{fig-flux-analysis}, right). 
However, similar to hemophilia alone, without amplification the thrombin signal is quickly extinguished by the combined action of ATIII and APC generated by thrombin.

\subsection*{Sensitivity analysis of the ensemble of reduced order coagulation models.}
We conducted a variance based global sensitivity analysis to better understand which parameters were controlling the performance of the reduced order coagulation model. 
In particular, we calculated the sensitivity of two performance metrics, the time to maximum thrombin and the thrombin exposure (area under the thrombin curve) to
changes in the model parameters (Fig. ZZ).

@@

\section*{Discussion}
Controlling hemophilia and more generally hemorrhage has been a driving force to understand the molecular basis of blood coagulation.
Replacement therapy improves bleeding times for most hemophilia patients. However, repeated fVIII/fIX administration can lead to neutralizing inhibitors. 
Alternatively, rFVIIa has been used as a universal hemostatic agent to initiate clotting with and without fVIII or fIX inhibitors \cite{Hedner:2008rm}.
Despite its wide adoption, the mechanism and the utility of rFVIIa in trauma-associated hemorrhage remains controversial \cite{Duchesne:2008hw}. 
In this study, we formulated a model of the human coagulation cascade and studied thrombin formation in normal and rFVIIa treated and untreated
hemophilic plasma. The model described 193 proteins or protein complexes connected by 301 interactions. 
The coagulation network architecture was based on literature and a previous mathematical model from our laboratory \cite{VARNER2007}.
Simulations of thrombin dynamics were independently validated using blood drawn from patients with coronary artery disease (reported here) and previous TF/FVIIa and FVIIa 
\emph{in-vitro} studies \cite{BUTENAS2002}.
While the human coagulation cascade is perhaps an ideal model system to develop network analysis tools, not all the model parameters were identifiable given the current training data.
Instead, an \emph{ensemble} of models (N = 437) was estimated using nine cell-based coagulation training sets \cite{ALLEN2007}.
We used this family of consistent coagulation models in all simulations instead of a single best-fit but uncertain model.
Additionally, using sensitivity analysis on the ensemble of models, we identified 
structurally sensitive components of the coagulation architecture as a function of condition.
We also demonstrated that the ensemble robustly constrained model predictions of independent validation sets, despite having many poorly constrained parameters.
Thus, the model ensemble displayed \emph{sloppy} behavior similar to that observed by Sethna and coworkers for other signal transduction networks \cite{Gutenkunst2007}.  
By studying a family of models, we perhaps partially addressed the uncertainty stemming from the many poorly characterized model parameters.
However, many other factors could influence our results. For example, missing structural interactions or biophysical factors such as blood flow could play a large role.
Thus, while the results presented here may be a valuable first step, more studies are required.

@@

>>>>

The model ensemble predicted a direct correlation between thrombin generation and rFVIIa addition in hemophilia (Fig. \ref{fig-rFVII}).
In the current model, we cannot distinguish between different initiation sources, e.g., TF/FVIIa and rFVIIa, as we have only a single lumped initiation source (trigger).
Thus, we simulated the addition of rFVIIa in hemophilia by removing fVIII and fIX from the model, and modulating the initial level of trigger.
Simulations with a baseline level of trigger were consistent with the previous hemophilia simulations, where the only thrombin being produced was from initiation (Fig. \ref{fig-rFVII}, 1$\times$~trigger).
As we increased the strength of the trigger event, the thrombin peak time and the maximum value of thrombin increased (Fig. \ref{fig-rFVII}, 50$\times$~trigger). 
However, as the trigger strength increased, the thrombin generated quickly decayed.
For example, for the highest trigger strength simulated (50$\times$trigger), 95\% of the thrombin was gone by 20 min after initiation. 
We performed flux analysis to understand how the reduced order coagulation model balanced initiation, amplification and inhibition of thrombin formation for normal coagulation and hemophilia.util..
Analysis of the reaction flux through the reduced order network for thrombin generation in normal, hemophilia and rFVIIa-treated hemophilia identified three distinct operational modes (Fig. \ref{fig-flux-analysis}).
The reduced order network includes four lumped reactions, initiation, amplification, thrombin-induced APC generation and total thrombin inhibition (including both APC and ATIII action).
Directly after the addition of a trigger (e.g., TF/FVIIa or rFVIIa), the lumped initiation flux was the largest for all three cases.
However, within a few minutes enough thrombin was generated by the initiation mechanism to induce the amplification stage. 
During amplification, thrombin catalyzes its own formation and inhibition by generating activated protein C (APC), a potent inhibitor of the coagulation cascade.
For normal coagulation, amplification and thrombin inhibition are the dominate reactions by 6 min after initiation (Fig. \ref{fig-flux-analysis}, left). 
After 10 min, the dominate reaction has shifted to thrombin inhibition (both ATIII and APC action). In the current proof-of-principle model, 
APC inhibits upstream of amplification thus APC activity will slow the rate of thrombin formation directly from the initiation trigger. 
On the other hand, ATIII inhibits thrombin directly, as well as upstream coagulation factors.
In hemophilia (missing both fVIII and fIX), the amplification reaction does not occur and the only thrombin produced is from initiation (Fig. \ref{fig-flux-analysis}, center).
Initiation is quickly inhibited by APC, and the thrombin level stabilizes and eventually decays because of ATIII activity. 
Lastly, when 50$\times$rFVIIa is used to induce thrombin formation in hemophilia (absence of fVIII/fIX), 
initiation mechanisms dominate for up to 6 min following initiation (Fig. \ref{fig-flux-analysis}, right). 
However, similar to hemophilia alone, without amplification the thrombin signal is quickly extinguished by the combined action of ATIII and APC generated by thrombin.

\subsection*{Sensitivity analysis of the ensemble of reduced order coagulation models.}
We conducted a global sensitivity analysis to determine which parameters controlled the performance of the reduced order coagulation model. 
We calculated the sensitivity of two performance metrics, the time to maximum thrombin and the thrombin exposure (area under the thrombin curve) with respect to
changes in parameter values (Fig. \ref{fig-sensitivity}).

@@

\section*{Discussion}
Controlling hemorrhage has been a driving force to understand the molecular basis of blood coagulation.
Replacement therapy improves bleeding times for most hemophilia patients. However, repeated fVIII/fIX administration can lead to neutralizing inhibitors. 
Alternatively, rFVIIa has been used as a universal hemostatic agent to initiate clotting with and without fVIII or fIX inhibitors \cite{Hedner:2008rm}.
Despite its wide adoption, the mechanism and the utility of rFVIIa in trauma-associated hemorrhage remains controversial \cite{Duchesne:2008hw}. 
In this study, we formulated a model of the human coagulation cascade and studied thrombin formation in normal and rFVIIa treated and untreated
hemophilic plasma. The model described 193 proteins or protein complexes connected by 301 interactions. 
The coagulation network architecture was based on literature and a previous mathematical model from our laboratory \cite{VARNER2007}.
Simulations of thrombin dynamics were independently validated using blood drawn from patients with coronary artery disease (reported here) and previous TF/FVIIa and FVIIa 
\emph{in-vitro} studies \cite{BUTENAS2002}.
While the human coagulation cascade is perhaps an ideal model system to develop network analysis tools, not all the model parameters were identifiable given the current training data.
Instead, an \emph{ensemble} of models (N = 437) was estimated using nine cell-based coagulation training sets \cite{ALLEN2007}.
We used this family of consistent coagulation models in all simulations instead of a single best-fit but uncertain model.
Additionally, using sensitivity analysis on the ensemble of models, we identified 
structurally sensitive components of the coagulation architecture as a function of condition.
We also demonstrated that the ensemble robustly constrained model predictions of independent validation sets, despite having many poorly constrained parameters.
Thus, the model ensemble displayed \emph{sloppy} behavior similar to that observed by Sethna and coworkers for other signal transduction networks \cite{Gutenkunst2007}.  
By studying a family of models, we perhaps partially addressed the uncertainty stemming from the many poorly characterized model parameters.
However, many other factors could influence our results. For example, missing structural interactions or biophysical factors such as blood flow could play a large role.
Thus, while the results presented here may be a valuable first step, more studies are required.

@@

\begin{figure}
\centering
\includegraphics[width=1.0\textwidth]{./figs/Figure-9-Sensitivity.pdf}
\caption{Global sensitivity of thrombin peak time and thrombin exposure. }\label{fig-sensitivity}
\end{figure}
